Identification of protoneutron star g-modes in gravitational-wave data \pcd{[PCD: alt. title]Inference of proto-neutron star properties from gravitational wave data in core-collapse supernovae.}

### 2020-09-11 introduction.tex
Done with section II

\pcd{In this work, we present a method to infer from the GW data alone, the time evolution of some properties or the PNS, namely a combination of its mass and radius. For this purpose we have developed an algorithm to  
extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convection). Next, we use the universal relations obtained by \cite{Torres:2019b} to infer the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$, being 
$M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

This paper is organised as follows. Section II describes the details of the 2D CCSN used. Section III focuses on the algorithm that extracts the time evolution of a combination of the mass and radius of the PNS corresponding to a g-mode. Section IV shows the performance of the data analysis method for different GW detectors. Finally, we discuss the results in section V.

>>>>

\pcd{In this work, we present a method to infer from the GW data alone, the time evolution of some properties or the PNS, namely a combination of its mass and radius. For this purpose we have developed an algorithm to  
extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convection). Next, we use the universal relations obtained by \cite{Torres:2019b}\pcd{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being 
$M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

This paper is organised as follows. Section II describes the details of the CCSN simulations used in the paper. Section III focuses on the algorithm that extracts the time evolution of a combination of the mass and radius of the PNS corresponding to a g-mode. Section IV shows the performance of the data analysis method for different GW detectors. Finally, we discuss the results in section V.

### 2020-09-11 simulations.tex
Done with section II

\section{Core collapse supernova simulations}
\label{sec:simulations}

@@

The algorithm proposed in the article does not require accurate
waveforms but relies on the evolution of the frequencies of
oscillations depending on the PNS mass and radius.  To parametrize
this dependence, we have considered spherically symmetric
\citep{Torres:2019a} and two-dimensional axisymmetric models of
stellar core collapse simulated with two codes, CoCoNuT
(one-dimensional models) and AENUS-ALCAR
\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
(one- and two-dimensional models).  CoCoNuT
\citep{Cerda-Duran__2008__AA__GRMHD-code} is a code for general
relativistic hydrodynamics coupled to the Fast Multigroup Transport
scheme \citep{Mueller_Janka_2015_FMT} providing an approximate
description of the emission and transport of neutrinos.  AENUS-ALCAR
\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
combines special relativistic (magneto-)hydrodynamics, a modified
Newtonian gravitational potential approximating the effects of general
relativity \citep{Marek_etal__2006__AA__TOV-potential}, and a spectral
two-moment neutrino transport solver
\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}.
We included the relevant reactions between matter and neutrinos of all
flavours, i.e., emission and absorption by nucleons and nuclei,
electron-positron pair annihilation, nucleonic bremsstrahlung, and
scattering off nucleons, nuclei, and electrons.

To tune the parameters of the algorithm, we use two sets of 25 models
in the range of initial stellar masses
$M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$ simulated with the two codes.
They were carried out using six equations of state (EOSs).  They are
complemented by a set of 7 two-dimensional axisymmetric models
consisting of stellar core collapse of five stars with zero-age
main-sequence masses of $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$
evolved through the hydrostatic phases by
\cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}.
We performed one simulation of each stellar model using the equation
of state of \cite{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS} with
an incompressibility of $K = 220 \, \mathrm{MeV}$ (LS220) and added
comparison simulations with the SFHo EOS
\cite{Steiner_et_al__2013__apj__Core-collapseSupernovaEquationsofStateBasedonNeutronStarObservations}
and the EOS of
\cite{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations}
(GShen) for the one with $M_{\mathrm{ZAMS}} = 15 \, \Msol$ (see Table
\ref{Tab:2dSimList} for a list of models).  To this set of
simulations, we add the waveform of a two-dimensional model used in
\cite{Torres:2019a}, denoted \texttt{s20S}.  It corresponds to a star
with the same initial mass, $M_{\mathrm{ZAMS}} = 20 \, \Msol$, as for
one of the other 7 axisymmetric simulations, but was taken from a
newer set of stellar-evolution models
\cite{Woosley_Heger__2007__physrep__Nucleosynthesisandremnantsinmassivestarsofsolarmetallicity}.
It was evolved with the SFHo EOS.

We mapped the pre-collapse state of the stars to a spherical
coordinate system with $n_r = 400$ zones in radial direction
distributed logarithmically with a minimum grid width of
$(\Delta r)_{\mathrm{min}} = 400 \, \mathrm{m}$ and an outer radius of
$r_{\mathrm{max}} = 8.3 \times 10^{9} \, \mathrm{cm}$ and
$n_{\theta} = 128$ equidistant cells in angular direction.  For the
neutrino energies, we used a logarithmic grid with $n_e = 10$ bins up
to $240 \, \mathrm{MeV}$.

@@

\begin{table}
  \centering
  \begin{tabular}{c|cc|ccc}
    \hline
    Simulation & $M_\mathrm{ZAMS} [\Msol]$ & EOS & $t_{\mathrm{f}} [\mathrm{s}]$&
    explosion & $M_{\mathrm{PNS}} [\Msol]$
    \\ 
    \hline
    \texttt{s11} & 11.2 & LS220 & 1.86 & $\times$ &  1.47
    \\ 
    \texttt{s15} & 15.0 & LS220 & 1.66 & $\times$ &  2.00
    \\ 
    \texttt{s15S} & 15.0 & SFHo & 1.75 & $\times$ &  2.02
    \\ 
    \texttt{s15G} & 15.0 & GShen & 0.97 & $\times$ &  1.86
    \\ 
    \texttt{s20} & 20.0 & LS220 & 1.53 & $\times$ & 1.75
    \\ 
    \texttt{s20S} & 20.0 & SFHo & 0.87 & $\times$ & 2.05
    \\ 
    \texttt{s25} & 25.0 & LS220 & 1.60 & $0.91$ &  2.33
    \\ 
    \texttt{s40} & 40.0 & LS220 & 1.70 & $1.52$ &  2.23
    \\ \hline
  \end{tabular}
  \caption{%%
    List of axisymmetric simulations.  We present the name of the models, the initial
    mass of the progenitors, and the EOS used, and the  final
    post-bounce time of the simulations.  For models which explode,
    we list the time at which the shock starts to expand in column
    ``explosion''; otherwise, a $\times$ sign is displayed.  The final
    column indicates the mass of the PNS at the end of the
    simulation.  
    %%
  }
  \label{Tab:2dSimList}
\end{table}

>>>>

% !TEX root = ccsn.tex
\section{Core collapse supernova simulations}
\label{sec:simulations}

@@

\pcd{Unless other methods used GW astronomy, the} algorithm proposed in \pcd{this work} does not require accurate
waveforms \pcd{in order to infer the properties of the PNS.} \pcd{Instead, it relies on the evolution of the frequency of
oscillation of some particular modes, as seen in the GW spectrum. The frequency of these modes depends, in a universal way, 
on the surface gravity of the PNS ($r=M_{\rm PNS}/R_{\rm PNS}^2$), in the sense that if at a given time we observe GW emission
at a certain frequency $f$ we can determine univocally the value of the surface gravity, within a certain error, regardless of the details
of the numerical simulation. In this work we use two sets of simulations: i) The {\it model set}, composed by 1D simulations, which is used 
to build the universal relation (model), $r(f)$, linking the ratio $r$ with the observed frequency $f$, and ii) the {\it test set}, composed by 2D simulations, 
for which we know both the GW signal and the evolution of the ratio, $r (t)$, and that is used to test performance of the algorithm.}

\pcd{We have used two different numerical codes in our numerical simulations.} 
%CoCoNuT
%(one-dimensional models) and AENUS-ALCAR
%\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
%(one- and two-dimensional models).  
CoCoNuT
\citep{Dimmelmeier:2002,Dimmelmeier:2005} is a code for general
relativistic hydrodynamics coupled to the Fast Multigroup Transport
scheme \citep{Mueller_Janka_2015_FMT} providing an approximate
description of the emission and transport of neutrinos.  AENUS-ALCAR
\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
combines special relativistic (magneto-)hydrodynamics, a modified
Newtonian gravitational potential approximating the effects of general
relativity \citep{Marek_etal__2006__AA__TOV-potential}, and a spectral
two-moment neutrino transport solver
\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}.
We included the relevant reactions between matter and neutrinos of all
flavours, i.e., emission and absorption by nucleons and nuclei,
electron-positron pair annihilation, nucleonic bremsstrahlung, and
scattering off nucleons, nuclei, and electrons.

\pcd{For the {\it model set}, we use the $25$ spherically symmetric (1D) simulations of \citep{Torres:2019a}
including progenitors with zero-age main sequence (ZAMS) masses in the range 
$M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$. The set contains simulations using the
 two numerical codes and six different equations of state. Details can be found in
 \citep{Torres:2019a}. The reason to use one dimensional simulations for the model set
 is that the computational cost of those is significantly smaller than the cost of multidimensional
 simulations, so is easier to accumulate the statistics necessary to build a good model for $r(f)$.}

 \begin{table}
  \centering
  \begin{tabular}{c|ccc|ccc}
    \hline
    Model & $M_\mathrm{ZAMS} $  & progenitor& EOS & $t_{\mathrm{f}}$& $t_{\rm explosion}$ & $M_{\mathrm{PNS, f}}$\\
    name&  $[\Msol]$ & model & &  $[\mathrm{s}]$& & $[\Msol]$ 
    \\ 
    \hline
    \texttt{s11} & 11.2 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& LS220 & 1.86 & $\times$ &  1.47 
    \\ 
    \texttt{s15} & 15.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& LS220 & 1.66 & $\times$ &  2.00 
        \\ 
    \texttt{s15S} & 15.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& SFHo & 1.75 & $\times$ &  2.02 
        \\ 
    \texttt{s15G} & 15.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& GShen & 0.97 & $\times$ &  1.86
         \\ 
    \texttt{s20} & 20.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& LS220 & 1.53 & $\times$ & 1.75 
        \\ 
    \texttt{s20S} & 20.0 & \cite{Woosley_Heger__2007__physrep__Nucleosynthesisandremnantsinmassivestarsofsolarmetallicity} & SFHo & 0.87 & $\times$ & 2.05 
    \\ 
    \texttt{s25} & 25.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& LS220 & 1.60 & $0.91$ &  2.33 
        \\ 
    \texttt{s40} & 40.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& LS220 & 1.70 & $1.52$ &  2.23 
        \\ \hline
  \end{tabular}
  \caption{%%
    List of axisymmetric simulations \pcd{used for the {\it test set}}.  
    \pcd{The last three columns show, the post-bounce time at the end of the
    simulation, the one at the onset of the explosion (non exploding models marked
    with $\times$), and the PNS mass at the end of the simulation.}
    %%
  }
  \label{Tab:2dSimList}
\end{table}

\pcd{For the {\it test set}, we use $8$ axisymmetric (2D) simulations using the AENUS-ALCAR code
(see Table~\ref{Tab:2dSimList} for a list of models).
$7$ of these simulations use a selection of progenitors with masses in the range} $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$
 evolved through the hydrostatic phases by
\cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}.
We performed one simulation of each stellar model using the equation
of state of \cite{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS} with
an incompressibility of $K = 220 \, \mathrm{MeV}$ (LS220) and added
comparison simulations with the SFHo EOS
\cite{Steiner_et_al__2013__apj__Core-collapseSupernovaEquationsofStateBasedonNeutronStarObservations}
and the EOS of
\cite{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations}
(GShen) for the \pcd{progenitor} with $M_{\mathrm{ZAMS}} = 15 \, \Msol$.  To this set of
simulations, we add the waveform of a two-dimensional model used in
\cite{Torres:2019a}, denoted \texttt{s20S}.  It corresponds to a star
with the same initial mass, $M_{\mathrm{ZAMS}} = 20 \, \Msol$, as for
one of the other 7 axisymmetric simulations, but was taken from a
newer set of stellar-evolution models
\cite{Woosley_Heger__2007__physrep__Nucleosynthesisandremnantsinmassivestarsofsolarmetallicity}.
It was evolved with the SFHo EOS.

\pcd{For all the simulations,} we mapped the pre-collapse state of the stars to a spherical
coordinate system with $n_r = 400$ zones in radial direction
distributed logarithmically with a minimum grid width of
$(\Delta r)_{\mathrm{min}} = 400 \, \mathrm{m}$ and an outer radius of
$r_{\mathrm{max}} = 8.3 \times 10^{9} \, \mathrm{cm}$ and
$n_{\theta} = 128$ equidistant cells in angular direction.  For the
neutrino energies, we used a logarithmic grid with $n_e = 10$ bins up
to $240 \, \mathrm{MeV}$.

@@
